---
title: "Fishbrain"
output: html_document
date: "2023-06-09"
---

Load packages
```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
```

Read in Fishbrain data
```{r}
fishbrain <- read_csv("fishbrain.csv")
View(fishbrain)
```

Filter to north of Pt. Conception

We decided to do this because the regulations in Southern district are complex.
```{r}
fishbrain_nc <- fishbrain %>% 
  filter(latitude > 34.45)
View(fishbrain_nc)
```

Add districts

I created a new column that separated the data into the four CDFW districts based on the latitude of each data point.
```{r}
fish_dis <- fishbrain_nc %>% 
  mutate(district = case_when(latitude > 34.45 & latitude < 37.18 ~ 'Central',
                              latitude >= 37.18 & latitude < 38.91 ~ 'Bay Area',
                              latitude > 38.91 & latitude < 40.44 ~ 'Wine',
                              latitude > 40.44 ~ 'Redwood')
  )
View(fish_dis)
```

Extract day, month, year columns

The caught_at_local_time column was a little difficult to work with on its own, so we decided to separate it into year, month, and day.
```{r}
fish_mdy <- fish_dis %>% 
  separate(caught_at_local_time, c('year', 'month', 'day'))
View(fish_mdy)
```

Species by year

I first converted the units of fish length from m to cm so the values wouldn't be fractions. I then converted fish weight from kg to g for the same reason. I then decided to look at the relationship between species and year. This table shows the mean length and weight of each of the 12 species with lengths and weights recorded from 2018-2023.
```{r}
#Convert length to cm
fish_cm <- fish_mdy %>% 
  mutate(length_in_cm = length_in_m * 100) %>% 
  drop_na()

#Convert weight to g
fish_g <- fish_cm %>% 
  mutate(weight_in_g = weight_in_kg * 1000)
View(fish_g)

#Summary table
species_yr <- fish_g %>% 
  group_by(year, common_name) %>% 
  summarize_if(is.numeric, mean) %>% 
  select(year, common_name, length_in_cm, weight_in_g)
View(species_yr)

#NOTE: Gopher rockfish, kelp rockfish, painted greenling, and treefish were excluded from this analysis because their length values were all NA.
```

Species by district

I then decided to look at the relationship between species and district. This table shows the mean weight and length of the previously shown 12 species by district.
```{r}
species_dis <- fish_g %>% 
  group_by(district, common_name) %>% 
  summarize_if(is.numeric, mean) %>% 
  select(district, common_name, length_in_cm, weight_in_g)
View(species_dis)

#NOTE: Gopher rockfish, kelp rockfish, painted greenling, and treefish were excluded from this analysis because their length values were all NA.
```

Plot #1: length by species

This first plot shows the relationship between species and length. 
```{r}
#Summary table
species_lgt <- fish_cm %>% 
  group_by(common_name) %>% 
  summarize(mean_length_cm = mean(length_in_cm)) 
View(species_lgt)
  
#Graph
species_lgt %>% 
  ggplot(aes(common_name, mean_length_cm)) + 
  geom_point() + 
  ylim(0, 50) +
  labs(x = "Species", y = "Mean Length (cm)") + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

#NOTE: Gopher rockfish, kelp rockfish, painted greenling, and treefish were excluded from this analysis because their length values were all NA.
```

Plot #2: length by district

This second plot shows the mean length for each district.
```{r}
#Summary table
dis_lgt <- fish_cm %>% 
  group_by(district) %>% 
  summarize(mean_length_cm = mean(length_in_cm))
View(dis_lgt)

#Graph
dis_lgt %>% 
  ggplot(aes(district, mean_length_cm)) +
  geom_point() + 
  ylim(0, 50) +
  labs(x = "District", y = "Mean Length (cm)")

#NOTE: Gopher rockfish, kelp rockfish, painted greenling, and treefish were excluded from this analysis because their length values were all NA.
```

Number of fish by month

We decided to look at the number of fish caught for each month. Using the assumption that one fish was caught for each data point, I first created a summary table that created a new column called fish_count which counted the number of rows for each month. I then created separate tables for each species showing how many of those species had been caught each month. 
```{r}
#All species
fish_month <- fish_mdy %>% 
  group_by(month) %>% 
  mutate(count = n()) %>% 
  summarize(fish_count = median(count))
View(fish_month)

#Black and yellow rockfish
fish_month_by <- fish_mdy %>% 
  filter(common_name == "Black and yellow rockfish") %>% 
  group_by(month) %>% 
  mutate(count = n()) %>% 
  summarize(fish_count_by = median(count))
View(fish_month_by)

#Black rockfish
fish_month_bla <- fish_mdy %>% 
  filter(common_name == "Black rockfish") %>% 
  group_by(month) %>% 
  mutate(count = n()) %>% 
  summarize(fish_count_bla = median(count))
View(fish_month_bla)

#Blue rockfish
fish_month_blu <- fish_mdy %>% 
  filter(common_name == "Blue rockfish") %>% 
  group_by(month) %>% 
  mutate(count = n()) %>% 
  summarize(fish_count_blu = median(count))
View(fish_month_blu)

#Brown rockfish
fish_month_br <- fish_mdy %>% 
  filter(common_name == "Brown rockfish") %>% 
  group_by(month) %>% 
  mutate(count = n()) %>% 
  summarize(fish_count_br = median(count))
View(fish_month_br)

#Cabezon
fish_month_ca <- fish_mdy %>% 
  filter(common_name == "Cabezon") %>% 
  group_by(month) %>% 
  mutate(count = n()) %>% 
  summarize(fish_count_ca = median(count))
View(fish_month_ca)

#China rockfish
fish_month_ch <- fish_mdy %>% 
  filter(common_name == "China rockfish") %>% 
  group_by(month) %>% 
  mutate(count = n()) %>% 
  summarize(fish_count_ch = median(count))
View(fish_month_ch)

#Copper rockfish
fish_month_co <- fish_mdy %>% 
  filter(common_name == "Copper rockfish") %>% 
  group_by(month) %>% 
  mutate(count = n()) %>% 
  summarize(fish_count_co = median(count))
View(fish_month_co)

#Gopher rockfish
fish_month_go <- fish_mdy %>% 
  filter(common_name == "Gopher rockfish") %>% 
  group_by(month) %>% 
  mutate(count = n()) %>% 
  summarize(fish_count_go = median(count))
View(fish_month_go)

#Grass rockfish
fish_month_gr <- fish_mdy %>% 
  filter(common_name == "Grass rockfish") %>% 
  group_by(month) %>% 
  mutate(count = n()) %>% 
  summarize(fish_count_gr = median(count))
View(fish_month_gr)

#Kelp greenling
fish_month_kg <- fish_mdy %>% 
  filter(common_name == "Kelp greenling") %>% 
  group_by(month) %>% 
  mutate(count = n()) %>% 
  summarize(fish_count_kg = median(count))
View(fish_month_kg)

#Kelp rockfish
fish_month_kr <- fish_mdy %>% 
  filter(common_name == "Kelp rockfish") %>% 
  group_by(month) %>% 
  mutate(count = n()) %>% 
  summarize(fish_count_kr = median(count))
View(fish_month_kr)

#Olive rockfish
fish_month_ol <- fish_mdy %>% 
  filter(common_name == "Olive rockfish") %>% 
  group_by(month) %>% 
  mutate(count = n()) %>% 
  summarize(fish_count_ol = median(count))
View(fish_month_ol)

#Painted greenling
fish_month_pg <- fish_mdy %>% 
  filter(common_name == "Painted greenling") %>% 
  group_by(month) %>% 
  mutate(count = n()) %>% 
  summarize(fish_count_pg = median(count))
View(fish_month_pg)

#Quillback rockfish
fish_month_qb <- fish_mdy %>% 
  filter(common_name == "Quillback rockfish") %>% 
  group_by(month) %>% 
  mutate(count = n()) %>% 
  summarize(fish_count_qb = median(count))
View(fish_month_qb)

#Rock greenling
fish_month_rg <- fish_mdy %>% 
  filter(common_name == "Rock greenling") %>% 
  group_by(month) %>% 
  mutate(count = n()) %>% 
  summarize(fish_count_rg = median(count))
View(fish_month_rg)

#Treefish
fish_month_tr <- fish_mdy %>% 
  filter(common_name == "Treefish") %>% 
  group_by(month) %>% 
  mutate(count = n()) %>% 
  summarize(fish_count_tr = median(count))
View(fish_month_tr)
```

Weight-length relationship

We decided to explore the weight-length relationship of all species. I decided to group by district since year isn't as reliable of a predictor because of increased user base throughout the years.
```{r}
#Summary table
dis_lgt_wt <- fish_g %>% 
  group_by(district) %>% 
  summarize(mean_length_cm = mean(length_in_cm), mean_weight_g = mean(weight_in_g))
View(dis_lgt_wt)
```

Read in RecFIN data
```{r}
recfin <- read_csv("recfin_fish.csv")
View(recfin)
```

Comparing Fishbrain to RecFIN data

This is the beginning of my comparisons between the Fishbrain and RecFIN data. I first decided to compare number of fish caught between the two. There are some discrepancies, particularly for the Wine district (Mendocino County).
```{r}
#Number of fish caught by district - Fishbrain
fish_count_dis <- fish_dis %>% 
  group_by(district) %>% 
  mutate(count = n()) %>% 
  summarize(fish_count = median(count))
View(fish_count_dis)

#Number of fish caught by district - RecFIN
recfin_dis <- recfin %>% 
  group_by(DISTRICT_NAME) %>% 
  summarize(mean_total_mortality_num = mean(SUM_TOTAL_MORTALITY_NUM))
View(recfin_dis)

#Total mortality in g - RecFIN
recfin_g <- recfin %>% 
  mutate(sum_total_mortality_g = SUM_TOTAL_MORTALITY_MT * 1000000) %>% 
  group_by(DISTRICT_NAME) %>% 
  summarize(mean_total_mortality_g = mean(sum_total_mortality_g))
View(recfin_g)
```

